---
title: "Housing Data Ocerview"
output: html_notebook
---

Installing packages

```{r}
library(readxl)
install.packages("SmartEDA")
library("SmartEDA")
install.packages("dplyr")
library("dplyr")
load.libraries <- c('data.table', 'testthat', 'gridExtra', 'corrplot', 'GGally', 'ggplot2', 'e1071', 'dplyr')

install.lib <- load.libraries[!load.libraries %in% installed.packages()]
for(libs in install.lib) install.packages(libs, dependences = TRUE)
sapply(load.libraries, require, character = TRUE)
```
```{r}
install.packages("mice")
```

Importing the data
```{r}
AmesHousing <- read_excel("/Users/chelseafitzpatrick/Downloads/AmesHousing.xls")
View(AmesHousing)
```

Home prices graph
```{r}
hist(AmesHousing$SalePrice, xlab = "Sale Price", main = paste("Histogram of Sale Price"))
```

Glimpse at data
```{r}
glimpse(AmesHousing)
```

Changing variable names 
```{r}
names(AmesHousing) <- tolower(names(AmesHousing))

names(AmesHousing) <- gsub(" ", "_", names(AmesHousing))
```

Reformating character variables to numeric and condensing categories
```{r}
table(AmesHousing$street)

AmesHousing$street_num[AmesHousing$street == "Pave"] <- 1
AmesHousing$street_num[AmesHousing$street != "Pave"] <- 0

table(AmesHousing$street_num)
```
```{r}
table(AmesHousing$alley)

AmesHousing$al[AmesHousing$alley == "NA"] <- 1
AmesHousing$al[AmesHousing$alley != "NA"] <- 0

table(AmesHousing$al)
```
```{r}
table(AmesHousing$lot_shape)

AmesHousing$lot[AmesHousing$lot_shape == "Reg"] <- 1
AmesHousing$lot[AmesHousing$lot_shape != "Reg"] <- 0

table(AmesHousing$lot)
```
```{r}
table(AmesHousing$ms_zoning)


AmesHousing$zone[AmesHousing$ms_zoning == c("RH", "RL", "RM")] <- 1
AmesHousing$zone[AmesHousing$ms_zoning != c("RH", "RL", "RM")] <- 0

table(AmesHousing$zone)

```
```{r}
table(AmesHousing$land_contour)


AmesHousing$land[AmesHousing$land_contour == "Lvl"] <- 1
AmesHousing$land[AmesHousing$land_contour != "Lvl"] <- 0

table(AmesHousing$land)
```
```{r}
table(AmesHousing$lot_config)


AmesHousing$lot_location[AmesHousing$lot_config == "Inside"] <- 1
AmesHousing$lot_location[AmesHousing$lot_config != "Inside"] <- 0

table(AmesHousing$lot_location)
```
```{r}
table(AmesHousing$land_slope)


AmesHousing$land_sl[AmesHousing$land_slope == "Gtl"] <- 1
AmesHousing$land_sl[AmesHousing$land_slope != "Gtl"] <- 0

table(AmesHousing$land_sl)
```
```{r}
table(AmesHousing$roof_style)


AmesHousing$roof[AmesHousing$roof_style == "Gable"] <- 1
AmesHousing$roof[AmesHousing$roof_style != "Gable"] <- 0

table(AmesHousing$roof)
```
```{r}
table(AmesHousing$roof_matl)


AmesHousing$roof_m[AmesHousing$roof_matl == "CompShg"] <- 1
AmesHousing$roof_m[AmesHousing$roof_matl != "CompShg"] <- 0

table(AmesHousing$roof_m)
```
```{r}
table(AmesHousing$heating)


AmesHousing$heat[AmesHousing$heating == "GasA"] <- 1
AmesHousing$heat[AmesHousing$heating != "GasA"] <- 0

table(AmesHousing$heat)
```
```{r}
table(AmesHousing$sale_condition)


AmesHousing$sale_cond[AmesHousing$sale_condition == "Normal"] <- 1
AmesHousing$sale_cond[AmesHousing$sale_condition != "Normal"] <- 0

table(AmesHousing$sale_cond)
```
```{r}
table(AmesHousing$condition_1)


AmesHousing$condition1[AmesHousing$condition_1 == "Norm"] <- 1
AmesHousing$condition1[AmesHousing$condition_1 != "Norm"] <- 0

table(AmesHousing$condition1)
```
```{r}
table(AmesHousing$bldg_type)


AmesHousing$build[AmesHousing$bldg_type == "1Fam"] <- 1
AmesHousing$build[AmesHousing$bldg_type != "1Fam"] <- 0

table(AmesHousing$build)
```
```{r}
table(AmesHousing$exter_qual)

AmesHousing$extertior[AmesHousing$exter_qual == "TA"] <- 1
AmesHousing$extertior[AmesHousing$exter_qual != "TA"] <- 0

table(AmesHousing$extertior)
```
```{r}
table(AmesHousing$exter_cond)

AmesHousing$extertior_cond[AmesHousing$exter_cond == "TA"] <- 1
AmesHousing$extertior_cond[AmesHousing$exter_cond != "TA"] <- 0

table(AmesHousing$extertior_cond)
```
```{r}
table(AmesHousing$central_air)

AmesHousing$CA[AmesHousing$central_air == "Y"] <- 1
AmesHousing$CA[AmesHousing$central_air != "Y"] <- 0

table(AmesHousing$CA)
```
```{r}
table(AmesHousing$electrical)

AmesHousing$elec[AmesHousing$electrical == "SBrkr"] <- 1
AmesHousing$elec[AmesHousing$electrical != "SBrkr"] <- 0

table(AmesHousing$elec)
```
```{r}
table(AmesHousing$functional)

AmesHousing$func[AmesHousing$functional == "Typ"] <- 1
AmesHousing$func[AmesHousing$functional != "Typ"] <- 0

table(AmesHousing$func)
```
Deleting old columns that were recategorized / or categories that could not be adjusted
```{r}
new_df <- subset(AmesHousing, select=-c(sale_condition, heating, roof_matl, street, lot_shape, ms_zoning, land_contour, lot_config, land_slope, roof_style, lot_frontage, alley, utilities, neighborhood, condition_1, condition_2, bldg_type, house_style, exterior_1st, exterior_2nd, mas_vnr_type, exter_qual, exter_cond, foundation, bsmt_qual, bsmt_cond, bsmt_exposure, bsmtfin_type_1, bsmtfin_type_2, heating_qc, central_air, electrical, kitchen_qual, functional, fireplace_qu, garage_type, garage_finish, garage_qual, garage_cond, paved_drive, pool_qc, fence, misc_feature, sale_type, order))
```
Means
```{r}
ExpData(data=new_df,type=2, fun = c("mean", "median", "var"))
```

Does the dataset still contain any missing values
```{r}
any(is.na(new_df))
sum(is.na(new_df))
```

Determining which variables include missing values 
```{r}
na.cols = which(colSums(is.na(new_df)) > 0)

sort(colSums(sapply(new_df[na.cols], is.na)), decreasing = TRUE)
```
Remove any missing cases from the dataset
```{r}
x <- new_df[complete.cases(new_df), ]
str(x)
```


removing outliers
```{r}
outlierMinMax <- quantile(x$saleprice, c(.01, .99))
Ames_clean <- x[ x$saleprice > outlierMinMax[1] & x$saleprice < outlierMinMax[2],  ]
```

Glimpse new Ames
```{r}
glimpse(x$saleprice)
glimpse(Ames_clean$saleprice)
```
Train/test split
```{r}
set.seed(1)
# create partition for train data
train_index <- createDataPartition(y = df$saleprice, p = 0.7, list = FALSE)
# index train and test data
train <- df[train_index, ]
test <- df[-train_index, ]
write.csv(Ames_clean, "cleaned_housing.csv")
write.csv(AmesHousing, "ames_housing.csv")
write.csv(train, "train_housing.csv")
write.csv(test, "test_housing.csv")
```
Train/test split
```{r}
set.seed(44)
# create partition for train data
train_index <- createDataPartition(y = df$saleprice, p = 0.7, list = FALSE)
# index train and test data
train2 <- df[train_index, ]
test2 <- df[-train_index, ]
write.csv(train2, "train_housing2.csv")
write.csv(test2, "test_housing2.csv")
```
Train/test split
```{r}
set.seed(15)
# create partition for train data
train_index <- createDataPartition(y = df$saleprice, p = 0.7, list = FALSE)
# index train and test data
train3 <- df[train_index, ]
test3 <- df[-train_index, ]
write.csv(train3, "train_housing3.csv")
write.csv(test3, "test_housing3.csv")
```
Train/test split
```{r}
set.seed(50)
# create partition for train data
train_index <- createDataPartition(y = df$saleprice, p = 0.7, list = FALSE)
# index train and test data
train4 <- df[train_index, ]
test4 <- df[-train_index, ]
write.csv(train4, "train_housing4.csv")
write.csv(test4, "test_housing4.csv")
```
Train/test split
```{r}
set.seed(3)
# create partition for train data
train_index <- createDataPartition(y = df$saleprice, p = 0.7, list = FALSE)
# index train and test data
train5 <- df[train_index, ]
test5 <- df[-train_index, ]
write.csv(train5, "train_housing5.csv")
write.csv(test5, "test_housing5.csv")
```